---
title: "STAT 8330 FALL 2015 ASSIGNMENT 8"
author: "Peng Shao"
date: "November 24, 2015"
output: 
        pdf_document:
                includes:
                        in_header: mystyles8.sty
---
```{r prerun, include=FALSE, cache=TRUE}
setwd("~/Documents/git/DataAnalysis3")
rm(list = ls())
library(plyr)
library(maps)
library(mapdata)
library(maptools)  
library(scales)
library(ggplot2)
library(mapproj)
library(ggmap)
library(ggbiplot)
library(ggfortify)
library(caret)
library(kernlab)
library(RDRToolbox)
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
        library(grid)
        
        # Make a list from the ... arguments and plotlist
        plots <- c(list(...), plotlist)
        
        numPlots = length(plots)
        
        # If layout is NULL, then use 'cols' to determine layout
        if (is.null(layout)) {
                # Make the panel
                # ncol: Number of columns of plots
                # nrow: Number of rows needed, calculated from # of cols
                layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                                 ncol = cols, nrow = ceiling(numPlots/cols))
        }
        
        if (numPlots==1) {
                print(plots[[1]])
                
        } else {
                # Set up the page
                grid.newpage()
                pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
                
                # Make each plot, in the correct location
                for (i in 1:numPlots) {
                        # Get the i,j matrix positions of the regions that contain this subplot
                        matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
                        
                        print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                                        layout.pos.col = matchidx$col))
                }
        }
}




set.seed(1)
ssts <- read.table("./SST011970_032003.dat.txt")
locs <- read.table("./SSTlonlat.dat.txt")
landmasks <- read.table("./SSTlandmask.dat.txt")
indexes <- read.table("./Nino34index.dat")
cnames <- expand.grid(month.abb, 1970:2002)
cnames <- mdply(cnames, 'paste', sep = "")
cnames <- c(cnames[, 3], "Jan2003", "Feb2003", "Mar2003")
colnames(ssts) <- cnames
names(locs) <- c("lon", "lat")
names(landmasks) <- "land"
rownames(indexes) <- cnames
keep <- (landmasks == 0)
temps <- data.frame(ave = apply(ssts, 1, mean), 
                    sd = apply(ssts, 1, sd))
p1 <- ggplot(data = locs[keep, ], 
             aes(x = lon, y = lat, color = temps$ave[keep]))
p1 <- p1 + geom_point() + 
        scale_colour_gradient(name = "Mean Temp.", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("Mean SST from Different Location") +
        theme(plot.title = element_text(lineheight=1, 
                                        face="bold", 
                                        size = 12))
p2 <- ggplot(data = locs[keep, ], 
             aes(x = lon, y = lat, color = temps$sd[keep]))
p2 <- p2 + 
        geom_point() + 
        scale_colour_gradient(name = "Std. Dev.", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("Standard Deviation of SST from Different Locations") +
        theme(plot.title = element_text(lineheight=1, 
                                        face="bold", 
                                        size = 12))
multiplot(p1, p2, cols=1)


pr.input <- t(ssts[keep, ])
pr.out.nonsd <- prcomp(pr.input)
pr.out.sd <- prcomp(pr.input, scale. = TRUE)
pr.out.nonsd.var <- pr.out.nonsd$sdev^2
pr.out.sd.var <- pr.out.sd$sdev^2
loadingmap1 <- ggplot(data = locs[keep, ], 
                      aes(x = lon, y = lat, color = pr.out.nonsd$rotation[, 1]))
loadingmap1 <- loadingmap1 + 
        geom_point() + 
        scale_colour_gradient(name = "Loading", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("First Component Loading Maps") +
        theme(plot.title = element_text(lineheight=1, face="bold"))
loadingmap2 <- ggplot(data = locs[keep, ], 
                      aes(x = lon, y = lat, color = pr.out.nonsd$rotation[, 2]))
loadingmap2 <- loadingmap2 + 
        geom_point() + 
        scale_colour_gradient(name = "Loading", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("Second Component Loading Maps") +
        theme(plot.title = element_text(lineheight=1, face="bold"))
pc1series <- ggplot(data.frame(pc1 = pr.out.nonsd$x[, 1]), aes(x = 1:399, y = pc1))
pc1series <- pc1series + geom_line() +
        xlab("Month") +
        ylab("PC1 Scores") +
        ggtitle("First Component Time Series") +
        theme(plot.title = element_text(lineheight=1, face="bold"))
pc2series <- ggplot(data.frame(pc2 = pr.out.nonsd$x[, 2]), aes(x = 1:399, y = pc2))
pc2series <- pc2series + geom_line() + 
        xlab("Month") +
        ylab("PC2 Scores") +
        ggtitle("Second Component Time Series") +
        theme(plot.title = element_text(lineheight=1, face="bold"))
multiplot(loadingmap1, pc1series, loadingmap2, pc2series, cols=2)



set.seed(3)
Z <- pr.out.nonsd$x[, 1:4]
km.out=kmeans(Z , 3, nstart=20)
km.plot <- ggplot(data = data.frame(Z, class = factor(km.out$cluster)), 
                  aes(x = PC1, y = PC2, colour = class))
km.plot <- km.plot + geom_point()
cl.plot <- ggplot(data = data.frame(Z, class = factor(indexes[, 1])), 
                  aes(x = PC1, y = PC2, colour = class))
cl.plot <- cl.plot + geom_point()
multiplot(km.plot, cl.plot, cols=1)
confusionMatrix(km.out$cluster, indexes[, 1]+2)


hc.average <- hclust(dist(Z), method="average")
hc.complete <- hclust(dist(Z), method="complete")
hc.average.cluster <- c(2, 1, 3)[cutree(hc.average, 3)]
hc.complete.cluster <- c(2, 1, 3)[cutree(hc.complete, 3)]
confusionMatrix(hc.average.cluster, indexes[, 1]+2)
confusionMatrix(hc.complete.cluster, indexes[, 1]+2)

sigma.grid <- c(0.01, 0.005, 0.001, 0.00001)
kpca.out <- c()
kpca.p <- list()
for (i in 1:length(sigma.grid)){
        kpca.out <- c(kpca.out, kpca( ~ ., data = data.frame(pr.input), 
                                      kpar = list(sigma = sigma.grid[i])))
        kpca.p[[i]] <- ggplot(data = data.frame(kpca.out[[i]]@rotated[, 1:2], 
                                                class = factor(indexes[, 1])), 
                              aes(x = X1, y = X2, colour = class)) + 
                geom_point() + 
                ggtitle(i)
}
multiplot(plotlist = kpca.p, cols = 2)
ggplot(data = (temp = data.frame(t(ssts[sample(1:nrow(ssts), 2), ]),class = factor(indexes[, 1]))), aes(x = temp[, 1], y = temp[,2], colour = class))+geom_point()+xlab(names(temp)[1])+ylab(names(temp)[2])


four_result <- data.frame(matrix(NA, ncol = 4, nrow = 0))
names(four_result) <- c("time", "kernel_width", "PC", "Scores")
for (i in 1:length(sigma.grid)){
        if (exists("d")){
                rm(d)
        }
        if (exists("temp")){
                rm(temp)
        }
        d = data.frame(time = 1:nrow(kpca.out[[i]]@rotated[, 1:2]),
                       kernel_width = rep(sigma.grid[i], nrow(kpca.out[[i]]@rotated[, 1:2])),
                       kpca.out[[i]]@rotated[, 1:2])
        temp <- reshape(data = d, 
                        direction = "long", 
                        varying = list(names(d)[3:4]), 
                        v.names = "Scores",
                        idvar = c("time", "kernel_width"),
                        timevar = "PC", 
                        times = c("PC1", "PC2"))
        rownames(temp) <- NULL
        four_result <- rbind(four_result, temp)
}

multi.time.plot <- ggplot(data = four_result, 
                          aes(x = time, y = Scores)) +
        geom_line() +
        facet_grid(kernel_width ~ PC)


k = c(5, 10, 20, 40)
isomap.out <- list()
LLE.out <- list()
for (i in 1:length(k)){
        isomap.out[[i]] <- Isomap(data = pr.input, k = k[i])
        LLE.out[[i]] <- Isomap(data = pr.input, k = k[i])
}
isomap.out <- lapply(isomap.out, 
                     data.frame, 
                     class = factor(indexes[, 1]))
LLE.out <- lapply(LLE.out, 
                  data.frame, 
                  class = factor(indexes[, 1]))
isomap.p <- list()
for (i in 1:length(k)){
        isomap.p[[i]] <- ggplot(data = isomap.out[[i]], 
                                aes(x = dim2.1, y = dim2.2, colour = class)) + 
                geom_point()
}
multiplot(plotlist = isomap.p, cols = 2)
LLE.p <- list()
for (i in 1:length(k)){
        LLE.p[[i]] <- ggplot(data = LLE.out[[i]], 
                             aes(x = dim2.1, y = dim2.2, colour = class)) + 
                geom_point()
}
multiplot(plotlist = LLE.p, cols = 2)
iso.lle.time.series <- rbind(cbind(time = 1:nrow(isomap.out[[4]]), 
                                   method = rep("Isomap", 
                                                nrow(isomap.out[[4]])), 
                                   isomap.out[[4]][, -3]), 
                             cbind(time = 1:nrow(LLE.out[[4]]), 
                                   method = rep("LLE", 
                                                nrow(LLE.out[[4]])), 
                                   LLE.out[[4]][, -3])
)
iso.lle.time.series <- reshape(data = iso.lle.time.series, 
                               direction = "long", 
                               varying = list(names(iso.lle.time.series)[3:4]), 
                               v.names = "Scores", 
                               idvar = c("time", "method"),
                               timevar = "Dim", 
                               times = c("First Dim", "Second Dim")
)
rownames(iso.lle.time.series) <- NULL
iso.lle.time.plot <- ggplot(data = iso.lle.time.series, 
                            aes(x = time, y = Scores)) +
        geom_line() + 
        facet_grid(method ~ Dim)

pp.1 <- isomap.p[[4]]
pp.2 <- LLE.p[[4]]
pp.3 <- kpca.p[[3]]
pp.4 <- ggplot(data = data.frame(pr.out.nonsd$x[, 1:2], 
                                 class = factor(indexes[, 1])), 
               aes(x = PC1, y = PC2, colour = class)) + 
        geom_point()
multiplot(pp.1, pp.3, pp.2, pp.4, cols = 2)
```
```{r, include=FALSE}
library(plyr)
library(maps)
library(mapdata)
library(maptools)  
library(scales)
library(ggplot2)
library(mapproj)
library(ggmap)
library(ggbiplot)
library(ggfortify)
library(caret)
library(kernlab)
library(RDRToolbox)
set.seed(1)
```

$\blacktriangleright$ \textbf{1.\quad Solution.}

From the two plots, the mean temperatures of the area does not show so much differerces. The only abnormal phenomenon is that the the loacations which have relatively low temperature are around the equator and eastern Pacific Ocean, where people usually think the weather might be hot. Moreover, these locations also have very high variation, while the temperatures of other locations of Pacific Ocean are much more stable.

```{r, echo=FALSE}
multiplot(p1, p2, cols=1)
```


$\blacktriangleright$ \textbf{2.\quad Solution.}

The variance accounted for by the first 4 principal components when the data are not standardized and when they are standardized are
```{r}
pr.out.nonsd.var[1:4]
pr.out.sd.var[1:4]
```

And the percentages of them are
```{r}
pr.out.nonsd.var[1:4]/sum(pr.out.nonsd.var)
pr.out.sd.var[1:4]/sum(pr.out.sd.var)
```



Both map plots show that the locations around the equator and eastern Pacific Ocean have high loading, which means that they play very important roles in the first principle component and the second principle component. This is in accordance with the result in question 1. And from the time series plots, there are several unusual low temperatures in the trend of PC1 and several unusual high temperatures in the trend of PC2. To sum up, based on PCA, the first two key features which enfluence the "Tropical Pacific Sea Surface Temperature (SST)", are the unusual low temperatures and usual high temperatures of the area between 180 W ~ 90 W, 5 S ~ 5 N(about Nino Region 3), which may be the well-known La Ni$\tilde{\text{n}}$a and El Ni$\tilde{\text{n}}$o.



```{r, echo=FALSE, cache=TRUE}
loadingmap1 <- ggplot(data = locs[keep, ], 
                      aes(x = lon, y = lat, color = -pr.out.nonsd$rotation[, 1]))
loadingmap1 <- loadingmap1 + 
        geom_point() + 
        scale_colour_gradient(name = "Loading", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("First Component Loading Maps") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 12))
loadingmap2 <- ggplot(data = locs[keep, ], 
                      aes(x = lon, y = lat, color = pr.out.nonsd$rotation[, 2]))
loadingmap2 <- loadingmap2 + 
        geom_point() + 
        scale_colour_gradient(name = "Loading", 
                              low = "darkblue", 
                              high = "yellow") +
        ylab("Latitude") +
        xlab("Longitude") + 
        ggtitle("Second Component Loading Maps") +
        theme(plot.title = element_text(lineheight=1, 
                                        face="bold", 
                                        size = 12))
pc1series <- ggplot(data.frame(pc1 = pr.out.nonsd$x[, 1]), aes(x = 1:399, y = pc1))
pc1series <- pc1series + geom_line() +
        xlab("Month") +
        ylab("PC1 Scores") +
        ggtitle("First Component Time Series") +
        theme(plot.title = element_text(lineheight=1, 
                                        face="bold", 
                                        size = 12))
pc2series <- ggplot(data.frame(pc2 = pr.out.nonsd$x[, 2]), aes(x = 1:399, y = pc2))
pc2series <- pc2series + geom_line() + 
        xlab("Month") +
        ylab("PC2 Scores") +
        ggtitle("Second Component Time Series") +
        theme(plot.title = element_text(lineheight=1, 
                                        face="bold", 
                                        size = 12))
multiplot(loadingmap1, pc1series, loadingmap2, pc2series, cols=2)
```

$\blacktriangleright$ \textbf{3.\quad Solution.}

To be clear, I use the No.3 random seed. The reason of my choice is the classification results of K-means cluster can be the same order with the Nion 3.4 Index, which means class 1 v.s La Ni$\tilde{\text{n}}$a, class 2 v.s Normal, and class 3 v.s El Ni$\tilde{\text{n}}$o. 

```{r, cache=TRUE, echo=FALSE, fig.height=4}
ts.plot <- ggplot(data = data.frame(Month = 1:399, 
                                    class = km.out$cluster), 
                  aes(x = Month, y = class))
ts.plot <- ts.plot + geom_line() + 
        ggtitle("Class Time Series") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 12))
km.plot <- ggplot(data = data.frame(Z, class = factor(km.out$cluster)), 
                  aes(x = PC1, y = PC2, colour = class))
km.plot <- km.plot + geom_point() + 
        ggtitle("K-means Cluster Classification") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 12))
cl.plot <- ggplot(data = data.frame(Z, class = factor(indexes[, 1])), 
                  aes(x = PC1, y = PC2, colour = class))
cl.plot <- cl.plot + geom_point() +
        ggtitle("True Classification") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 12))
multiplot(ts.plot, km.plot, cl.plot, 
          layout = matrix(c(1,1,2,3), nrow=2, byrow=TRUE))
```


From the classification plots, we can see that most data are correctly classified, so the K-means cluster based on PCA is fairly good. To see the performance more precisely, we can create the confusion matrix.

```{r}
confusionMatrix(km.out$cluster-2, indexes[, 1])$table
confusionMatrix(km.out$cluster-2, indexes[, 1])$overall["Accuracy"]
```


$\blacktriangleright$ \textbf{4.\quad Solution.}

For the average linkage methods
```{r}
confusionMatrix(hc.average.cluster-2, indexes[, 1])$table
confusionMatrix(hc.average.cluster-2, indexes[, 1])$overall["Accuracy"]
```
For the complete linkage methods
```{r}
confusionMatrix(hc.complete.cluster-2, indexes[, 1])$table
confusionMatrix(hc.complete.cluster-2, indexes[, 1])$overall["Accuracy"]
```


Both methods are worse than the K-means cluster, but the the complete linkage method is a little better than the other, not because the accuracy, but the high sensitivity, which make it possible to detect the anomalies more precisely.

$\blacktriangleright$ \textbf{5.\quad Solution.}



```{r, cache=TRUE, echo=FALSE}
multi.time.plot + ggtitle("Time Series Plots of PC1 and PC2 with Different Width") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 10))
```

```{r, cache=FALSE, echo=FALSE}
kpca.out <- c()
kpca.p <- list()
for (i in 1:length(sigma.grid)){
        kpca.out <- c(kpca.out, kpca( ~ ., data = data.frame(pr.input), 
                                      kpar = list(sigma = sigma.grid[i])))
        kpca.p[[i]] <- ggplot(data = data.frame(kpca.out[[i]]@rotated[, 1:2], 
                                                class = factor(indexes[, 1])), 
                              aes(x = X1, y = X2, colour = class)) + 
                geom_point() + 
                xlab("PC1") + 
                ylab("PC2") +
                ggtitle(paste("Radial Basis kernel with sigma=", 
                              sigma.grid[i], sep = "")) +
                theme(plot.title = element_text(lineheight = 1, 
                                                face = "bold", 
                                                size = 10))
}
multiplot(plotlist = kpca.p, cols = 2)
```

```{r, eval=FALSE}
library(kernlab)
library(ggplot2)
sigma.grid <- c(0.01, 0.005, 0.001, 0.00001)
kpca.out <- c()
kpca.p <- list()
for (i in 1:length(sigma.grid)){
        kpca.out <- c(kpca.out, kpca( ~ ., data = data.frame(pr.input), 
                                      kpar = list(sigma = sigma.grid[i])))
        kpca.p[[i]] <- ggplot(data = data.frame(kpca.out[[i]]@rotated[, 1:2], 
                                                class = factor(indexes[, 1])), 
                              aes(x = X1, y = X2, colour = class)) + 
                geom_point() + 
                xlab("PC1") + 
                ylab("PC2") +
                ggtitle(paste("Radial Basis kernel with sigma=", 
                              sigma.grid[i], sep = "")) +
                theme(plot.title = element_text(lineheight = 1, 
                                                face = "bold", 
                                                size = 10))
}
multiplot(plotlist = kpca.p, cols = 2)
four_result <- data.frame(matrix(NA, ncol = 4, nrow = 0))
names(four_result) <- c("time", "kernel_width", "PC", "Scores")
for (i in 1:length(sigma.grid)){
        if (exists("d")){
                rm(d)
        }
        if (exists("temp")){
                rm(temp)
        }
        d = data.frame(time = 1:nrow(kpca.out[[i]]@rotated[, 1:2]),
                       kernel_width = rep(sigma.grid[i], nrow(kpca.out[[i]]@rotated[, 1:2])),
                       kpca.out[[i]]@rotated[, 1:2])
        temp <- reshape(data = d, 
                        direction = "long", 
                        varying = list(names(d)[3:4]), 
                        v.names = "Scores",
                        idvar = c("time", "kernel_width"),
                        timevar = "PC", 
                        times = c("PC1", "PC2"))
        rownames(temp) <- NULL
        four_result <- rbind(four_result, temp)
}

multi.time.plot <- ggplot(data = four_result, 
                          aes(x = time, y = Scores)) +
        geom_line() +
        facet_grid(kernel_width ~ PC)
```



$\blacktriangleright$ \textbf{6.\quad Solution.}

```{r, cache=TRUE, echo=FALSE}
iso.lle.time.plot + ggtitle("Time Series Plots of ISOMAP and LLE Using 40  Neighbors")
```


```{r, cache=TRUE, echo=FALSE}
pp.1 <- isomap.p[[4]] + ggtitle("Isomap with 40 Neighbors") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 12)) +
        xlab("Dim 1") +
        ylab("Dim 2")
pp.2 <- LLE.p[[4]] + ggtitle("LLE with 40 Neighbors") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 10))
pp.3 <- kpca.p[[3]] + ggtitle(paste("Kernel PCA with Width=",
                                    sigma.grid[3], 
                                    sep = "")) +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 10)) +
        xlab("Dim 1") +
        ylab("Dim 2")
pp.4 <- ggplot(data = data.frame(pr.out.nonsd$x[, 1:2], 
                                 class = factor(indexes[, 1])), 
               aes(x = PC1, y = PC2, colour = class)) + 
        geom_point() + 
        ggtitle("PCA") +
        theme(plot.title = element_text(lineheight = 1, 
                                        face = "bold", 
                                        size = 10))
multiplot(pp.1, pp.3, pp.2, pp.4, cols = 2)
```




```{r, eval=FALSE}
library(RDRToolbox)
k = c(5, 10, 20, 40)
isomap.out <- list()
LLE.out <- list()
for (i in 1:length(k)){
        isomap.out[[i]] <- Isomap(data = pr.input, k = k[i])
        LLE.out[[i]] <- Isomap(data = pr.input, k = k[i])
}
isomap.out <- lapply(isomap.out, 
                     data.frame, 
                     class = factor(indexes[, 1]))
LLE.out <- lapply(LLE.out, 
                  data.frame, 
                  class = factor(indexes[, 1]))
isomap.p <- list()
for (i in 1:length(k)){
        isomap.p[[i]] <- ggplot(data = isomap.out[[i]], 
                                aes(x = dim2.1, y = dim2.2, colour = class)) + 
                geom_point()
}
multiplot(plotlist = isomap.p, cols = 2)
LLE.p <- list()
for (i in 1:length(k)){
        LLE.p[[i]] <- ggplot(data = LLE.out[[i]], 
                             aes(x = dim2.1, y = dim2.2, colour = class)) + 
                geom_point()
}
multiplot(plotlist = LLE.p, cols = 2)
iso.lle.time.series <- rbind(cbind(time = 1:nrow(isomap.out[[4]]), 
                                   method = rep("Isomap", 
                                                nrow(isomap.out[[4]])), 
                                   isomap.out[[4]][, -3]), 
                             cbind(time = 1:nrow(LLE.out[[4]]), 
                                   method = rep("LLE", 
                                                nrow(LLE.out[[4]])), 
                                   LLE.out[[4]][, -3])
)
iso.lle.time.series <- reshape(data = iso.lle.time.series, 
                               direction = "long", 
                               varying = list(names(iso.lle.time.series)[3:4]), 
                               v.names = "Scores", 
                               idvar = c("time", "method"),
                               timevar = "Dim", 
                               times = c("First Dim", "Second Dim")
)
rownames(iso.lle.time.series) <- NULL
iso.lle.time.plot <- ggplot(data = iso.lle.time.series, 
                            aes(x = time, y = Scores)) +
        geom_line() + 
        facet_grid(method ~ Dim)

pp.1 <- isomap.p[[4]]
pp.2 <- LLE.p[[4]]
pp.3 <- kpca.p[[3]]
pp.4 <- ggplot(data = data.frame(pr.out.nonsd$x[, 1:2], 
                                 class = factor(indexes[, 1])), 
               aes(x = PC1, y = PC2, colour = class)) + 
        geom_point()
multiplot(pp.1, pp.3, pp.2, pp.4, cols = 2)
```









